####Very rough - certainly MANY issues exist.. 

---
title: "full_parental.19apr18.Rmd"
author: "MacManes"
date: "4/19/2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Set up env


```{r import libs, cache=TRUE, echo =FALSE}
library(edgeR)
library(tximport)
library(readr)
library(locfit)
library(statmod)
library(gplots)
library(topGO)
library(Rgraphviz)
library(beanplot)
library(knitr)
```

## Tximport
```{r tximport, cache=TRUE}
#dir <- setwd('/Volumes/4TB_1/')
dir <- setwd('/Users/macmnaes/Dropbox')
tx2gene <- read.table(file.path(dir, "RockDove_working/parental/parental1.tximport"), header=T)
all_samples <- read.table(file.path(dir, "RockDove_working/parental/19Apr18.list"), header=F)
all_files <- file.path(dir, "RockDove_working/mapping/files/parental", all_samples$V1, "quant.sf")
names(all_files) <- paste0(all_samples$V1)
txi.all <- tximport(all_files, type = "salmon", tx2gene = tx2gene, dropInfReps=TRUE)
cts <- txi.all$counts
normMat <- txi.all$length
normMat <- normMat/exp(rowMeans(log(normMat)))
o <- log(calcNormFactors(cts/normMat)) + log(colSums(cts/normMat))

```
## Add pheonotypes

```{r phenotype, cache=TRUE}

meta_final <- read_csv("RockDove_working/parental/Parental_meta_final.csv")
sex <- meta_final$sex
tissue <- meta_final$tissue
stage <- meta_final$stage
day <- meta_final$day
treatment <- meta_final$treatment

```

## Export file with pseudocounts
```{r export file, cache=TRUE}
write.table(cts, file="parentalcounts26Apr.txt", sep = "," , row.names = TRUE)
```

## Import file with gene names added
```{r import edited file, cache=TRUE}
parental_analysis <- read.csv("RockDove_working/parental/parentalcounts28Apr.txt", row.names="Name")
```

## Make groups
```{r make groups, cache=TRUE}
group <- factor(paste(sex,stage,tissue,day,treatment,sep="."))
final_matrix <- cbind(meta_final, group=group)

## make sure that I have the right groups assigned to the right samples
parentalobject$samples[ which(parentalobject$samples$group == "female.lay.pit.D1.normal"), ]

```

## Do EdgeR analysis

```{r do edgeR analysis, cache=TRUE}
parentalobject <- DGEList(counts=parental_analysis[,4:990], genes=parental_analysis[,1:3], group=final_matrix$group)

cpms <- cpm(parentalobject)
keep_genes <- rowSums(cpms >= 1) >= 10

## Filter the DGEList
dge <- parentalobject[keep_genes, ]

parentaldesign <- model.matrix(~ group )
colnames(parentaldesign) <- levels(group)
parentalobject <- calcNormFactors(parentalobject)
parentalobject <- estimateCommonDisp(parentalobject)
parentalobject <- estimateTagwiseDisp(parentalobject)
parentalobject <- estimateDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMCommonDisp(parentalobject, parentaldesign, verbose=TRUE)
parentalobject <- estimateGLMTrendedDisp(parentalobject, parentaldesign)
parentalobject <- estimateGLMTagwiseDisp(parentalobject, parentaldesign)
fit <- glmFit( parentalobject, parentaldesign, robust=T)
treatres <- glmTreat(fit, lfc = 1)
```

## Make contrasts

```{r make contrasts, cache=TRUE}
my.contrasts <- makeContrasts(
	         FPL_H = female.lay.pit.D1.normal
	         - female.hatch.pit.D18.normal,
	         HI_MF = female.inc.hypo.D17.normal
	         - male.inc.hypo.D17.normal,
	         PI_MF = female.inc.pit.D17.normal
	         - male.inc.pit.D17.normal,	         
	         HH_MF = female.hatch.hypo.D18.normal
	         - male.hatch.hypo.D18.normal,
	         PH_MF = female.hatch.pit.D18.normal
	         - male.hatch.pit.D18.normal,		         
	         FHL_H = female.lay.hypo.D1.normal
	         - female.hatch.hypo.D18.normal,
	         FPI_H = female.inc.pit.D17.normal
	         - female.hatch.pit.D18.normal,
	         FHI_H = female.inc.hypo.D17.normal
	         - female.hatch.hypo.D18.normal,
	         MPI_H = male.inc.pit.D17.normal
	         - male.hatch.pit.D18.normal,
	         MHI_H = male.inc.hypo.D17.normal
	         - male.hatch.hypo.D18.normal,
	         MGI_H = male.inc.gonad.D17.normal
	         - male.hatch.gonad.D18.normal,
	         FGI_H = male.inc.gonad.D17.normal
	         - male.hatch.gonad.D18.normal,
levels=parentaldesign)
```


## find DE Female pit lay to hatch

```{r, Female pit lay to hatch, cache=TRUE}

#n=10,10  

length(which(parentalobject$samples$group == "female.lay.pit.D1.normal"))
length(which(parentalobject$samples$group == "female.inc.pit.D17.normal"))
length(which(parentalobject$samples$group == "female.inc.hypo.D17.normal"))
length(which(parentalobject$samples$group == "female.hatch.pit.D18.normal"))

library(knitr)
cont <- "FPL_H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = 1), 
	adjust.method="fdr", p.value=0.01))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont]), n=5), digits=2, lfc = 1)

plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=1), main='Female pit hatch to lay', frame.plot=F)
```

## Beanplots for Female pit lay to hatch

```{r beanplots LH}
beanplotPIT_LH <- function(vals, genenames)
{
  gene <- parentalobject$pseudo.counts[grep(paste("\\b", vals, "\\b", sep=""), parentalobject$genes$entrezid), ]
  FPL <- log(gene[c(which(parentalobject$samples$group == "female.lay.pit.D1.normal"))]+1)
  FPH <- log(gene[c(which(parentalobject$samples$group == "female.hatch.pit.D18.normal"))]+1)
  MPL <- log(gene[c(which(parentalobject$samples$group == "male.lay.pit.D1.normal"))]+1)
  MPH <- log(gene[c(which(parentalobject$samples$group == "male.hatch.pit.D18.normal"))]+1)
  each=c(10,10,10,10)
  values <- c(FPL, FPH, MPL, MPH)
  treatment <- rep(c('Lay','Lay', 'Hatch', 'Hatch'), each)
  groupplot <- rep(c('female', 'male', 'female', 'male'), each)
  title="%s"
  require(beanplot)
  beanplot(values ~ treatment*groupplot, ll = 0, beanlinewd=0,
           main=do.call(sprintf, as.list(c(title, genenames))), side = "both", xlab="", ylab='logTPM',
           col = list("forestgreen","dodgerblue2","green3","deepskyblue"),
           axes=F)
  axis(2)
  arrows(1.5, median(FPL), 1.5, median(FPH), lwd=2, length=.1)
  arrows(1.4, median(MPL), 1.4, median(MPH), lwd=2, length=.1)
}
```

## find DE Female pit inc to hatch
```{r, Female pit inc to hatch, cache=TRUE}

#n=11,10  

length(which(parentalobject$samples$group == "female.inc.pit.D17.normal"))
length(which(parentalobject$samples$group == "female.hatch.pit.D18.normal"))



cont <- "FPI_H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), 
	adjust.method="fdr", p.value=0.05))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), n=39), digits=2)

plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=.5), main='Female pit inc to hatch', frame.plot=F)
topTags(glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), n=39)
```

## Beanplot Female pit inc to hatch
```{r beanplots IH}
beanplotPIT_IH <- function(vals, genenames)
{
  gene <- parentalobject$pseudo.counts[grep(paste("\\b", vals, "\\b", sep=""), parentalobject$genes$entrezid), ]
  FPI <- log(gene[c(which(parentalobject$samples$group == "female.inc.pit.D17.normal"))]+1)
  FPH <- log(gene[c(which(parentalobject$samples$group == "female.hatch.pit.D18.normal"))]+1)
  MPI <- log(gene[c(which(parentalobject$samples$group == "male.inc.pit.D17.normal"))]+1)
  MPH <- log(gene[c(which(parentalobject$samples$group == "male.hatch.pit.D18.normal"))]+1)
  each=c(length(FPI),length(FPH),length(MPI),length(MPH))
  values <- c(FPI, FPH, MPI, MPH)
  treatment <- rep(c('Inc','Hatch', 'Inc', 'Hatch'), each)
  groupplot <- rep(c('female', 'female', 'male', 'male'), each)
  title="%s"
  require(beanplot)
  beanplot(values ~ treatment*groupplot, ll = 0, beanlinewd=0,
           main=do.call(sprintf, as.list(c(title, genenames))), side = "both", xlab="", ylab='logTPM',
           col = list("forestgreen","dodgerblue2","green3","deepskyblue"),
           axes=F)
  axis(2)
  arrows(1.5, median(FPI), 1.5, median(FPH), lwd=2, length=.1)
  arrows(1.4, median(MPI), 1.4, median(MPH), lwd=2, length=.1)
}
```



## find DE Female hypo inc to hatch
```{r, Female pit inc to hatch, cache=TRUE}

#n=11,10  

length(which(parentalobject$samples$group == "female.inc.hypo.D17.normal"))
length(which(parentalobject$samples$group == "female.hatch.hypo.D18.normal"))



cont <- "FHI_H"
summary(decideTestsDGE(
	glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), 
	adjust.method="fdr", p.value=0.05))
kable(topTags(glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), n=39), digits=2)

plotMD(glmTreat(fit, contrast=my.contrasts[,cont], lfc=.5), main='Female hypo inc to hatch', frame.plot=F)
topTags(glmTreat(fit, contrast=my.contrasts[,cont], lfc = .5), n=39)
```
